The Effects of weather on maize yields: New evidence from Kenya

### 2018-12-18 writing/draft3/draft15.tex
writing...

To find the most suitable set of fixed effects, we adapted a stepwise selection procedure which minimized the Akaike's Information Criterion (AIC). The first step of the procedure was the estimation of the model with complete set of our weather measures in the fixed effects and with the random intercepts. After this we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors). We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects slightly improved the AIC. The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days). Furthermore, including them only improved AIC from $2080.877$ to $2076.439$ (these values differ from the AIC values presented
in table~\ref{ARMA} in the appendix because the final estimates compared in table~\ref{ARMA} were estimated using restricted maximum likelihood (REML), but the models were re-estimating using ML method for the comparison as REML comparisons are not meaningful). In addition, the signs of these variables were positive in contrary to our expectation. Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.

>>>>

To find the most suitable set of fixed effects, we adapted a stepwise selection procedure which minimized the Akaike's Information Criterion (AIC). The first step of the procedure was the estimation of the model with the complete set of our weather measures in the fixed effects and with the random intercepts. After this, we performed a search through the subsets of the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors). We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects would slightly improve the AIC. The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days). Furthermore, including the heatwave days and maximum daily temperature would only improve AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for the comparison\footnote{In general, REML is the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however REML criterion is not meaningful for comparison of the models with different fixed effects therefore the models need to be re-estimated by ML for comparison \cite{Zuur2009}.}). In addition, the coefficients of both heatwave days and maximum daily temperature were positive, which is opposite to what we expected. Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature and so we did not include them in our preferred model.
